---
title: "HW2"
author: "Santos Barrera"
date: "4/1/2020"
output:
  word_document: default
  html_document: default
  pdf_document: default
---


## Normalize the path:
```{r setup, include=FALSE} 
knitr::opts_chunk$set(echo = TRUE) 
knitr::opts_knit$set(root.dir=normalizePath('../../')) 
```

#1.  Version control

#2. Identify three traits of interest showing different levels of heterosis 

Traits selected according Flint-Garcia et al., 2009 data
PltYield: Plant yield (gr/plant)
LFLEN: Leaf Length (cm)
TSLLEN: Tassel length (cm) 

#3.Following the basic steps to estimate genetic variances

First, the linear model is written down

pijr = Î¼ + fi + Ej + eijr

pijr:phenotype value
Î¼: averall mean
fi: effect of the ith father
Ej: is the effect of the jth environment 
eijr: is the residual error ~ğ‘ğ‘œğ‘Ÿğ‘šğ‘ğ‘™(0,ğœ_ğ‘”^2)

Second, phenotypic variance is written in terms of components. 
ğœ^2P=ğœ^f+ğœ2E+ğœ^2e
ğœ2p=Phenotypic Variance
ğœ^2f=Genotpic Variance
ğœ^2E=Entironmental Variance
ğœ^2e=error Variance

Third, interpret the ANOVA table and variance components in terms of the covariances between relatives
I have a question:
For half-siblings (hybrids in Flint-Garcia et al., 2009 data) we have to consider the inbreeding coefficient (F), which is 1/2 for half-siblings, for the calculation of VA (additive variance). However, in your lecture and Rex Bernardo Book, we calculate it from the mean sq (MS) of the interaction gxe. What is the procedure for finding mean sq (MS) of the interaction gxe in R?. In this case do we treat the environment as random effect in the model? Because I couln't calculate this value I can't reporte the additive variance and therefore I can't report narrow sense heredability h^2

VA=4/1+FVprogeny


# Hybrids 
Plant yield (gr/plant)

```{r}
# library(lme4) when we account for randome effects 
library(Matrix)
f <- read.delim("data/journal.pone.0007433.s001.txt", header=TRUE)
# Convert missing data to NA
f[f=="."] <- NA

# four environments
table(f$Env)

f$INBRED <- as.factor(f$INBRED)
f$Env <- as.factor(f$Env)
# tricky part, be careful:
f$PltYield_Hyb  <- as.numeric(as.character((f$PltYield_Hyb)))
hist(f$PltYield_Hyb, col="grey", xlab="Plant yield (gr/plant)") 
fit_PYH <- lm(PltYield_Hyb ~ INBRED + Env, data=f)
#plot (fit_PYH)
anova(fit_PYH)
```
```{r}
Vg=  (2332 - 1210)/2 #only 2 locations for PltYield_Hyb (North Carolina and Missouri)
Ve = 1210

H2 = Vg/(Vg + Ve)
H2
```
# Hybrids 
Leaf Length (cm) 

```{r}
f$INBRED <- as.factor(f$INBRED)
f$Env <- as.factor(f$Env)
# tricky part, be careful:
f$LFLEN_Hyb  <- as.numeric(as.character((f$LFLEN_Hyb)))
hist(f$LFLEN_Hyb, col="grey", xlab="Plant Leaf Length (cm) ") 
fit_LFH <- lm(LFLEN_Hyb~ INBRED + Env, data=f)
#plot (fit_LFH)
anova(fit_LFH)
```

```{r}
Vg=  (16367 - 1430)/4
Ve = 1430 

H2 = Vg/(Vg + Ve)
H2
```

# Hybrids
Tassel length (cm) 

```{r}
f$INBRED <- as.factor(f$INBRED)
f$Env <- as.factor(f$Env)
# tricky part, be careful:
f$TSLLEN_Hyb  <- as.numeric(as.character((f$TSLLEN_Hyb)))
hist(f$TSLLEN_Hyb, col="grey", xlab="Plant Tassel length (cm) ") 
fit_TSLH <- lm(TSLLEN_Hyb~ INBRED + Env, data=f)
#plot (fit_TSLH)
anova(fit_TSLH)
```


```{r}
Vg=  (2185 - 405)/2 #only 2 locations for Tassel length (cm) (Florida and North Carolina)
Ve = 405 

H2 = Vg/(Vg + Ve)
H2
```


### Flint-Garcia et al., 2009 data
# Inbreed lines 
# Plant yield (gr/plant)

```{r}
f <- read.delim("data/journal.pone.0007433.s001.txt", header=TRUE)
# Convert missing data to NA
f[f=="."] <- NA

# four environments
table(f$Env)

f$INBRED <- as.factor(f$INBRED)
f$Env <- as.factor(f$Env)
# tricky part, be careful:
f$PltYield_Inbred  <- as.numeric(as.character((f$PltYield_Inbred)))
hist(f$PltYield_Inbred, col="grey", xlab="Plant yield (gr/plant)") 
fit_PYI <- lm(PltYield_Inbred ~ INBRED + Env, data=f)
#plot (fit_PYI)
anova(fit_PYI)
```
```{r}
Vg=  (549.08 - 358.99)/2 #only 2 locations for PltYield_Hyb (North Carolina and Missouri)
Ve = 358.99

H2 = Vg/(Vg + Ve)
H2
```


# Inbreed lines 
# Leaf Length (cm) 

```{r}
f$INBRED <- as.factor(f$INBRED)
f$Env <- as.factor(f$Env)
# tricky part, be careful:
f$LFLEN_Inbred <- as.numeric(as.character((f$LFLEN_Inbred)))
hist(f$LFLEN_Inbred, col="grey", xlab="Plant Leaf Length (cm) ") 
fit_LFI <- lm(LFLEN_Inbred~ INBRED + Env, data=f)
#plot (fit_LFI)
anova(fit_LFI)
```

```{r}
Vg=  (31302 - 2259)/4
Ve = 2259 

H2 = Vg/(Vg + Ve)
H2
```

# Inbreed lines 
# Tassel length (cm) 

```{r}
f$INBRED <- as.factor(f$INBRED)
f$Env <- as.factor(f$Env)
# tricky part, be careful:
f$TSLLEN_Inbred  <- as.numeric(as.character((f$TSLLEN_Inbred)))
hist(f$TSLLEN_Inbred, col="grey", xlab="Plant Tassel length (cm) ") 
fit_TSLI <- lm(TSLLEN_Inbred~ INBRED + Env, data=f)
#plot (fit_TSLI)
anova(fit_TSLI)
```


```{r}
Vg=  (3757 - 645)/2 #only 2 locations for PltYield_Hyb (Florida and North Carolina)
Ve = 645 

H2 = Vg/(Vg + Ve)
H2
```

#4. Visualization 

```{r}
library(ggplot2)
H2_hyb <- c(0.32, 0.72, 0.68)
H2_Inb <- c(0.20, 0.76, 0.70)
Trait <- c("Plant yield (gr/plant)", "Leaf Length (cm)", "Tassel length (cm) ")
dat <- data.frame(Trait = Trait ,H2_hyb = H2_hyb, H2_Inb = H2_Inb)
dat

ggplot(dat, aes(x= Trait, y=H2_hyb, fill=Trait)) +
  geom_bar(stat="summary", fun.y="heritability")
ggplot(dat, aes(x= Trait, y=H2_Inb, fill=Trait)) +
  geom_bar(stat="summary", fun.y="heritability")

```